### Executive Summary
Research, evaluate and apply machine learning to predict the manner in which participants did the exercise from Human Activity Recognition(HAR) data set from Groupware (See reference)

## Data source
groupware@les(http://groupware.les.inf.puc-rio.br/har) pis the data provider. The training data set is here (https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).
The test data is here(https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

``` {r, getdata, echo=TRUE, eval=FALSE}
train.url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
test.url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
download.file(train.url, paste('./data/','pml-training.csv',sep=''), method='wget')
download.file(test.url, paste('./data/','pml-testing.csv', sep=''),     method='wget')

```

## Data Cleansee
Eye balling dataset shows 'NA', "", and "#DIV/0!" are empty data. Lots of empty or mostly empty columns.
Also participants names, timestamps and windows are not from device, so remove those columnes.
``` {r, loaddata, echo=TRUE, eval=TRUE}
rm(list=ls())
traindata <- read.csv(file='data/pml-training.csv', header=T, stringsAsFactors=TRUE, na.strings=c("NA", "", "#DIV/0!"))
result<-traindata$classe
testdata <- read.csv(file='data/pml-testing.csv', header=T, stringsAsFactors=TRUE, na.strings=c("NA", "", "#DIV/0!"))
#
``` 
Find out which columns are empty(NA, null, DIV/0) or mostly empty and remove them. Replace empty entries with mean for "not so empty columns". Mostly empty columns as morethan 45% empty.
``` {r, cleansedata, echo=TRUE, eval=TRUE}
emptycols <- sapply(traindata, function(x)all(is.na(x)))
#remove all empty columns
traindata<-traindata[,-which(emptycols)]
#remove mostly empty columns
emptycol <- function(x) {
    numMissing <-length(which(is.na(x)))
    numCells <- length(x)
    numMissing/numCells
}
mostlyemptycols <- sapply(traindata, FUN=emptycol)
traindata<-traindata[,-which(mostlyemptycols>0.45)]
# remove names, timestamp etc
redudantcols <- c(which(colnames(traindata)=='user_name'),
which(colnames(traindata)=='raw_timestamp_part_1'),which(colnames(traindata)=='raw_timestamp_part_2'),which(colnames(traindata)=='cvtd_timestamp'),which(colnames(traindata)=='new_window'),which(colnames(traindata)=='num_window'))
traindata <- traindata[,-redudantcols]
# intersection of column names from train and test data
cmn <- intersect(names(traindata),names(testdata))
traindata <- traindata[,cmn]
testdata <- traindata[,cmn]
traindata <- traindata[,-c(which(colnames(traindata)=='X'))]
testdata <- traindata[,-c(which(colnames(testdata)=='X'))]
# add back the result
traindata$classe<-result
``` 
"Test data" set from groupware is for evaluation of final model. We partion the 
"Training data" into local train and local test data sets for buildig model.
We use a 70, 30 split.
``` {r, partitiondata, echo=TRUE, eval=TRUE}
library(caret)
set.seed(12312014)
trainingIndex  <- createDataPartition(traindata$classe, p=.70, list=FALSE)
training.train <- traindata[ trainingIndex,]
training.test  <- traindata[-trainingIndex,]
``` 
## Exploratory Analysis
Plot variables
``` {r, exploredata, echo=TRUE, eval=TRUE}


``` 
## Machine Learning Method

``` {r, machinelearn, echo=TRUE, eval=TRUE}
library(randomForest)
random.forest <- train(training.train[,-c(which(colnames(traindata)=='classe'))],
                       training.train$classe,
                       tuneGrid=data.frame(mtry=3),
                       trControl=trainControl(method="none")
                       )
``` 

## Diagnosis and Evaluation

``` {r, evalmachinelearn, echo=TRUE, eval=TRUE}
summary(random.forest)

# confusion matrix
confusionMatrix(predict(random.forest,
                        newdata=training.test[,-c(which(colnames(traindata)=='classe'))]
                ),
                training.test$classe
                )

plot(varImp(random.forest))
```
## Prediction on TestData

## Reference
* groupware@les (http://groupware.les.inf.puc-rio.br/har)
* Wearable Computing: Accelerometers’ Data
Classification of Body Postures and Movements (http://groupware.les.inf.puc-rio.br/public/papers/2012.Ugulino.WearableComputing.HAR.Classifier.RIBBON.pdf)
