### Executive Summary
Research, evaluate and apply machine learning to predict the manner in which participants did the exercise from Human Activity Recognition(HAR) data set from Groupware (See reference)

## Data source
groupware@les(http://groupware.les.inf.puc-rio.br/har) pis the data provider. The training data set is here (https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv).
The test data is here(https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv)

``` {r, getdata, echo=TRUE, eval=FALSE}
train.url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv'
test.url <- 'https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv'
download.file(train.url, paste('./data/','pml-training.csv',sep=''), method='wget')
download.file(test.url, paste('./data/','pml-testing.csv', sep=''),     method='wget')

```

## Data Cleanse
``` {r, loaddata, echo=TRUE, eval=TRUE}
rm(list=ls())
traindata <- read.csv(file='data/pml-training.csv', header=T, stringsAsFactors=FALSE)
``` 
Find out which columns are empty(NA, null, DIV/0) or mostly empty and remove them. Replace empty entries with mean for "not so empty columns"
``` {r, cleansedata, echo=TRUE, eval=TRUE}
nrow(is.na(traindata))
ncol(is.na(traindata))
mydata[!complete.cases(mydata),]
``` 


## Exploratory Analysis

``` {r, exploredata, echo=TRUE, eval=TRUE}

``` 
## Machine Learning Method

``` {r, machinelearn, echo=TRUE, eval=TRUE}
library(randomForest)
random.forest <- train(training.train[,-57],
                       training.train$classe,
                       tuneGrid=data.frame(mtry=3),
                       trControl=trainControl(method="none")
                       )
``` 

## Diagnosis and Evaluation

## Prediction on TestData

## Reference
* groupware@les (http://groupware.les.inf.puc-rio.br/har)
* Wearable Computing: Accelerometers’ Data
Classification of Body Postures and Movements (http://groupware.les.inf.puc-rio.br/public/papers/2012.Ugulino.WearableComputing.HAR.Classifier.RIBBON.pdf)
